---
title: 'Part 1: Simulation Exercise Instructions'
author: "Somaya AlGabry"
date: "23/10/2020"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---
In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda

```{r setup, include=TRUE}
set.seed(1)
lambda= 0.2
n= 40
mns = NULL
for (i in 1 : 1000) {
  mns= c(mns, mean( rexp(n, lambda))) 
}
```
# 1. Show the sample mean and compare it to the theoretical mean of the distribution.

``` {r mean, include= TRUE, echo= TRUE}

# Sample Mean:
SampleMean <- mean(mns)
SampleMean
# Theoretical Mean:
TheoreticalMean<- 1/lambda
TheoreticalMean
# Difference between theoretical and sample mean:
TheoreticalMean - SampleMean
# Visualization:
hist(mns, xlab = "mean", main = "Comparison between the Sample Mean and the theoretical mean ")
abline(v= TheoreticalMean, col= "blue" )
abline(v= SampleMean, col= "red")
```


# 2. show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

``` {r varianvce, include= TRUE, echo=TRUE}
# Sample Variance:
SampleVar <- var(mns)
SampleVar
# Theoretical Variance:
TheoreticalVar <- (1/lambda^2)/n
TheoreticalVar
# Difference between sample and theoretical variance:
TheoreticalVar - SampleVar

```

# 3. Show that the distribution is approximately normal.
I used three methods:

1. Histogram: there is a very little devition of the sample distribution (gray) from the theoretical bell curve of the normal distribution (red).

2. Box Plot: The symmetry indicates normal distribution.

3. QQ Plot: The quantiles of the variable are in line with the theoretical normal quantiles making a straight line, this is telling us we have a normal distribution.

``` {r distibution, include= TRUE, echo= TRUE}
#Histogram:

hist(mns, main = "Distribution of samples", xlab = "means", ylab = "density", probability = TRUE, breaks = n)
x <- seq(min(mns), max(mns), length=100)
curve(dnorm(x, mean=1/lambda, sd=(1/lambda/sqrt(n)) ), col="red", lwd=2, add=TRUE, yaxt="n")

#Box Plot:

boxplot(mns)

#QQ Plot:

qqnorm(mns)
qqline(mns)
```



